```{r}
library(readxl)
library(openxlsx)
library(tidyverse)
```

## Part 1 - creating main dataset

# 0.Modify excel file - add "Month" column with range 1-12

# 1.Read delivery and clients data

```{r}
clients_1_6 <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "Data_main_clients1-6")
clients_7_12 <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "Data_main_clients7-12")
delivery_all <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "Data_main_delivery")
```

# 2.Merge clients data with rbind(), remove excess columns

```{r}
clients_all <-   rbind(clients_1_6, ..... ) %>%
  select (-..... )
clients_all
```
# 3.Rename delivery table columns  
```{r}
delivery_all <- delivery_all %>%
  mutate ( Loadsheet_Delivered_fact_rcs =  Delivered_fact_rcs) %>%
 select ( c(-....., -....., -.....))
delivery_all
```



# 4.Join delivery and clients data

```{r}
clients_delivery_all <- merge(clients_all, delivery_all, by = '.....')
clients_delivery_all
```

# 5.Create a new table with number of stops

```{r}
number_of_stops <-  clients_delivery_all %>% group_by(loadsheet) 
number_of_stops
number_of_stops <-  summarise(number_of_stops, stops = n()) %>%
  mutate (stops)
number_of_stops
```

# 6.Add number of stops data to main table

```{r}
clients_delivery_all_stops <- merge(....)
clients_delivery_all_stops
```

# 7.Calculate KM to stop (add "km_to_stop" column)
```{r}


```

## Part 2 - calculating trdparty costs
# 8. Creating separate tables for trdparty and own shipments

```{r}
clients_delivery_all_stops <- mutate (clients_delivery_all_stops, Costs_trdparty = TrdpartyPrice/Loadsheet_Delivered_fact_rcs*Delivered_fact_rcs)

clients_delivery_all_stops_own <-

```

## Part 3 - creating additional rate tables

# 9.Get all cases(rcs), distance (kms)s and stops (stops) during month (own transport)

```{r}
month_group <-  ..... %>% group_by(Month) 

month_group <-  summarise(month_group, stops = n(), rcs = ..., kms = .... ) %>%
  mutate (stops,rcs, kms)

```

# 10. Get Repair_Maintenance data ("RM_Costs")

```{r}
r_m <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "Repair_Maintenance")
r_m <- pivot_longer (r_m, cols = 1:12, ..., ...)
r_m
```

# 11. Get Fuel data

```{r}
fuel <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "....")
fuel <- ..
fuel
```

# 12. Get Depriciation data

```{r}
depr_fd <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "....")
depr_fd <- ..
depr_fd
```

# 13. Get Fleet Labour_Benefits data

```{r}
l_b_fd <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "..")
l_b_fd <- ..
l_b_fd
```

# 14. Get Drivers_Labour_Benefits data

```{r}
l_b_drivers <- read_excel("E:/R/Cost_allocation.xlsx", sheet = "..")
l_b_drivers <- ..
l_b_drivers
```

# 15. Get month costs rate
```{r}
month_group <- merge(month_group, r_m,  by = 'Month')
month_group <- ..
month_group <- ..
month_group <- ..
month_group <- ..

```

# 16 Calculate month rates (example: R_M_Rate is R_M_Costs divided by total kilometers)

```{r}
month_group <- month_group %>%
  mutate(R_M_Rate =.., Fuel_Rate = .., .. = .., .. = .., .. = ..) %>%
  #Which columns should be selected?
  select (.. )

```

## Part 4. Add rates data to main table (own transport):

# 17. Join tables: main data table with aggregated month data

```{r}
clients_delivery_all_stops <- merge(..)
```

# 18. Calculating allocated costs (example: R_M_Alloc_Costs = R_M_Rate*km_to_stop)

```{r}
clients_delivery_all_stops_final <- clients_delivery_all_stops %>%
  mutate(..)

clients_delivery_all_stops_final 

```

#19. Removing excess data

```{r}
clients_delivery_all_stops_final2 <- clients_delivery_all_stops_final  %>%
  #This line changes all NA to zero values:
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(Total_Alloc_Costs =( R_M_Alloc_Costs + Fuel_Alloc_Costs + Depr_Alloc_Costs+
           FD_LB_Costs+ Drivers_LB_Costs+ Costs_trdparty ) ) %>%
 select (-c(Month, R_M_Rate,Fuel_Rate,Depr_Rate,FD_LB_Rate,Drivers_LB_Rate) )
clients_delivery_all_stops_final2

write.xlsx(clients_delivery_all_stops_final2, 'test_3.xlsx')

```
