---
title: "South India traveler's clusterization"
output:
  html_document:
    toc: yes
    toc_float: yes

---

```{r, message=FALSE, warning=FALSE}
library(tidyverse) 
library(NbClust)
library(factoextra)
library(cluster)
library(GGally)
```

# Dataset Information

The dataset [`traveler_reviews.csv`](https://archive.ics.uci.edu/ml/datasets/BuddyMove+Data+Set) was populated from destination reviews published by 249 reviewers of holidayiq.com till October 2014. Reviews falling in 6 categories among destinations across South India were considered and the count of reviews in each category for every reviewer (traveler) is captured.

Attribute Information:
- User Id: Unique user id
- Sports: Number of reviews on stadiums, sports complex, etc.
- Religious: Number of reviews on religious institutions
- Nature: Number of reviews on beach, lake, river, etc.
- Theatre: Number of reviews on theatres, exhibitions, etc.
- Shopping: Number of reviews on malls, shopping places, etc.
- Picnic: Number of reviews on parks, picnic spots, etc.


```{r, message=FALSE, warning=FALSE}
d <- read_csv('traveler_reviews.csv')
glimpse(d)
```

# Tasks

1) Prepare data for clustering. Analyze distributions of variables. Perform normalization.

```{r}

```

2) Choose the optimal number of clusters, using Elbow method, Silhouette method, Gap statistic method.

```{r}

```

3) Perform clustering using k-means algorithm. Visualize and interpret clusters.

```{r}

```

4) Evaluate clusterization by Silhouette plot.

```{r}

```

