---
title: "Loan defaults prediction"
author: "Morozova Y.A."
date: '21.04.2020'
output: html_document
---

```{r message = FALSE}
library(tidyverse)
library(mlr)
```

# Data structure

File `bankloan.csv` contains hypothetical data that concerns a bank's efforts to reduce the rate of loan defaults. The file contains financial and demographic information on 850 past and prospective customers. 

The first 700 cases are past customers, last 150 are prospective customers:

- age - Age in years  
- ed - Level of education (1 - "Did not complete high school", 2 - "High school degree", 3 - "Some college", 4 - "College degree")  
- employ - Years with current employer  
- address - Years at current address  
- income - Household income in thousands  
- debtinc - Debt to income ratio (x100)  
- creddebt - Credit card debt in thousands  
- default - Previously defaulted (0 - "No", 1 - "Yes")

```{r}
bankloan <- read_csv("bankloan.csv")
bankloan_past <- head(bankloan, 700)
bankloan_prosp <- tail(bankloan, 150)
```

# Tasks

1) Perform exploratory data analysis and choose predictors for loan default prediction model.

```{r}

```

2) Build logistic regression model for a loan default prediction and interpret estimated coefficients.

```{r}

```

3) Calculate loan default predictions for past customers using logistic regression model and compare it with actual data. Visualize results.

```{r}

```

4) Calculate loan default predictions for prospective customers using logistic regression model.