---
title: "Assignment: interval estimate for cars mileage"
author: "Gleb Zakhodyakin, postlogist@gmail.com"
date: "17.01.2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(digits = 4) #4 digits in output
```

```{r Load packages}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(readxl))
```

# Assignment
The cars dataset contains data on cars fuel efficiency (miles per gallon, mpg).
Compute mean and confidence intervals for given confidence levels for mpg.

Load the cars.xlsx dataset into the `cars` variable

```{r Load data}
cars <- ...
```

Review the structure of the dataset

```{r Review data}

```

Summarize the mpg variable using `summary()` function. How many missing values are there?

```{r Find missing data}

```

Remove observations with missing values in the `mpg` variable from the dataset.
Put the cleaned dataset into the same `cars` variable:

```{r Remove NAs}
cars <- ...
```


How many rows in the dataset are left after dropping observations with missing `mpg`?

```{r Number of rows}

```


Let's consider the dataset as the entire population of cars.
Compute the population mean and standard deviation

```{r Population parameters}
pop_mean <- ...
pop_sd <- ...
```


Set random seed using your favourite number

```{r Set seed}
set.seed(<your number here>)
```

Make a random sample of 9 cars and store it in cars9 variable

```{r Sampling}
cars9 <- ...
```

Compute sample mean and standard deviation

```{r Sample mean and SD}
sample_mean <- ...
sample_sd <- ...
```


Compute the standard error for the sample mean

```{r Standard error}
sample_se <- ...
```

Compute the half-width of 90% confidence interval based on sample mean and standard deviation

```{r Confidence interval half-width }
delta <- ...
```

Compute the upper and lower limits for a 90% confidence interval for a population mean
```{r Confidence limits}
lci <- ...
uci <- ...
```


Plot your mean and confidence interval using ggplot2:

```{r Plotting}

```


Post your results to this online form:  http://tinyurl.com/cars-mean-2017

 

