---
title: "Interval estimation of a population mean"
author: "Gleb Zakhodyakin, postlogist@gmail.com"
date: "17.01.2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(digits = 4) #4 digits in output
```

## Why care about confidence intervals?

A common research problem is making conclusions about a large **population** based on limited data. Such data are collected by randomly selecting a small **sample** from the population. Since sampling process involves randomness, there will be an error in estimates based on sample data - the **sampling error**. This simply means that if someone decides to explore the same population and studies another sample of the same size, he/she will get different results. 

Statistics provides several ways for estimating the uncertainty associated with estimates of population parameters based on sample data. First, there is a concept of a **standard error**. This statistic is computed from sample data and allows to estimate the variation of possible results based on samples of the same size. The standard error can be computed for any sample statistic.

Next, there is a concept of a **confidence interval**. The confidence interval is constructed around the estimate based on a sample. The interval's width allows to assess how close that estimate might be to the true value of a population parameter. A **confidence level** is associated with a confidence interval. This is a probability of getting the true value of a population parameter within the interval when doing a repeated estimation of that parameter based on samples of the same size.

In this example, we will illustrate the algorithm for constructing the **confidence interval for a population mean**.

## Generating population data
We will study a (fake) population of people with objective to estimate the average height. We will estimate the average height based on a small sample of just 9 people randomly selected from the entire population. To save us effort, we will generate the data on population height instead of going into the field and trying to measure it ourselves. The population data will be sampled using R's facilities. All in all, we are just trying to understand the statistical inference machinery...

First, we need to load packages and to generate population data.
Our population includes 10000 people. The height in population is normally distributed with mean 175 cm and standard deviation 10 cm.

```{r Load packages}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
```


```{r Generate population data}
set.seed(435)
height <- round(rnorm(10000, 175, 10)) 
```

Note the `set.seed()` function. It initializes the random number generator with a specific number. All the "random" numbers in a computer are actually only looking random, but in practice they are generated by a deterministic algorithm. Including the `set.seed()` in your script ensures reproducibility - we'll get the same results each time despite using "randomness".

The resulting distribution of heights in the population:

```{r Height histogram}
ggplot(data = tibble(height), aes(x = height)) +
  geom_histogram(binwidth = 2) +
  geom_vline(aes(xintercept = mean(height)), colour='red', linetype='dashed')+
  ggtitle('Height distribution in the population') +
  xlab('Height, cm')
```

## Sampling
Next, we draw a random sample from the population:

```{r Sampling}
# Sampling
set.seed(435)
N <- 9
height_sample <- sample(height, size = N)
```

**Note:** the `sample()` function will randomly select a given number of elements from a *vector*. It can't work with data frames, so if you need to sample a data frame, there's a trick with randomly sampling a vector with the row indices first, and then using the selected indices to extract the actual rows of the data frame:

```{r How to sample a data frame}
# A vector with row indices is sampled
my_df <- tibble(x = 1:10, y = 101:110)

set.seed(435)
selected <- sample(1:nrow(my_df), size = 5)

# These indices are used for extracting the selected rows:
my_df[selected, ] #rows in the sample
my_df[-selected, ] #rows not in the sample
```

## Interval estimation of the population mean

The first step is making *point estimates* of population parameters based on our sample.

```{r Sample statistics}
# The sample
height_sample

# Sample mean
sample_mean <- mean(height_sample)
paste("Sample mean:", sample_mean)

# Sample standard deviation
sample_sd <- sd(height_sample)
paste("Sample S.D.:", sample_sd)
```

Next, we will compute the standard error for the sample mean

```{r Standard Error for the sample mean}
# Standard error for the sample mean
sample_se <- sample_sd/sqrt(N)
paste("Std. error for the sample mean:", sample_se)
```

The next step is selecting the **confidence level** $P$ (usually .95) and computing the confidence interval half-width. Since we don't know the population standard deviation, we'll use the sample standard deviation for calculating the standard error. This will require using the Student's t-distribution instead of standard normal distribution to account for increased uncertainty.

```{r CI Half-Width}

P <- .95 # confidence level
alpha <- 1 - P

# 95% confidence interval half width
delta <- qt(1 - alpha / 2 , N - 1) * sample_se
paste("Confidence interval half-width:", round(delta, 1))
```

Last step is constructing the confidence interval around our estimate of population mean. For this, the half-width is added to/substracted from the sample mean. Upper and lower confidence limits are produced as a result.

```{r Confidence limits }
#  upper and lower confidence limits:
lcl <- sample_mean - delta
ucl <- sample_mean + delta

c(lcl, ucl)
```

## Comparing estimate and true value of the population mean

Finally, we compute the true population mean and compare our interval estimate to this value. Note that this is only possible since we already know the population.

```{r Compare estimate to actual value}

population_mean <- mean(height)
paste("Population mean:", round(population_mean, 1), 
      "Sample mean:", round(sample_mean, 1))

ggplot(data = tibble(population_mean, sample_mean, lcl, ucl),
       aes(x = 0)) +
  geom_point(aes(y = sample_mean)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), 
                col = 'red', width = 0.1) +
  labs(title = 'Interval estimate for a population mean',
       y = 'Mean Height, cm',
       x = NULL) +
  geom_hline(aes(yintercept = population_mean),
             col = 'red', linetype = 'dashed') +
  xlim(-1, 1)

```

## Computing confidence interval using a built-in funcion

There is a much simpler way to compute the confidence interval in R. We can use the built-in `t.test()` function

```{r Using t.test}
# A simpler way to get a confidence interval:
t.test(height_sample)

# Extracting the required part of the result:
r <- t.test(height_sample)
names(r)

tibble(by_hand = c(lcl, ucl), using_function = as.vector(r$conf.int))

```

There is also a simpler way of plotting the confidence interval using ggplot2. Actually, we don't need to compute the confience limits and can just use the built-in statistical function for computation.

```{r Confidence interval using ggplot2}
ggplot(data = tibble(height = height_sample)) + 
  stat_summary(aes(x = 0, y = height), 
               fun.data = 'mean_cl_normal', 
               colour = 'red') +
  labs(title = 'Interval estimate for a population mean',
       y = 'Mean Height, cm', x = NULL) +
  geom_hline(aes(yintercept = population_mean), 
             col = 'red', linetype = 'dashed')
```

